---
title: "OCN Productivity"
output: html_notebook
---

# Notes

```{r set up, echo=FALSE, message=FALSE, warning=FALSE}
# center drive working directory
workDir <- "//nwcfile/home/liermannma/CurrentWorrk/consulting/sept2018_sept2019/oregonCoho"

# local computer working directory
workDir <-"C:\\Users\\Martin.Liermann\\Documents\\projects\\oregonCoho"

dataDir <- paste(workDir,"data",sep="/")
library(coastalCohoSS)
```


```{r load the data, echo=FALSE, message=FALSE, warning=FALSE}
dat <- createJAGSdata(dataDir,"smolt",habType="miles")$fdat
hdat <- createJAGSdata(dataDir,"smolt",habType="miles")$hdat
```

Create a table with the necessary information.

```{r}
# create some new derived variables
tDat <- dat %>% left_join(hdat,by="Site") %>% 
  mutate(femaleSpawners=FemaleParentsWild+FemaleParentsHatchery,
         femaleSpawnersPerMile=femaleSpawners/miles,
         smoltPerMile=Smolts/miles,
         pHOS <- FemaleParentsHatchery/(FemaleParentsHatchery + FemaleParentsWild))

```

Now plot the data

```{r warning=FALSE, message=FALSE}
library(ggplot2)
ggplot(tDat,aes(x=femaleSpawners,y=Smolts)) +
  geom_point() +
  xlim(0,NA) +
  ylim(0,NA) +
  facet_wrap(.~Site,scales="free")
```

Now divide smolt and spawners by miles to place on a comparable scale.

```{r warning=FALSE, message=FALSE}
library(ggplot2)
ggplot(tDat,aes(x=femaleSpawnersPerMile,y=smoltPerMile)) +
  geom_point() +
  xlim(0,NA) +
  ylim(0,NA) +
  facet_wrap(.~Site,scales="free")
```

now divide by area

```{r warning=FALSE, message=FALSE}
library(ggplot2)
ggplot(tDat,aes(x=femaleSpawners/area,y=Smolts/area)) +
  geom_point() +
  xlim(0,NA) +
  ylim(0,NA) +
  facet_wrap(.~Site,scales="free")
```

Now fit spawner recruit functions.


```{r}
  hockeyStick <- function(S,a,b){
    pmin(log(a)+log(S),log(b))
  }

  bevHolt <- function(S,a,b){
    log(S/(1/a+S/b))
  }

  SS_HockeyStick <- function(p,S,R){
    sum((log(R)-hockeyStick(S,exp(p[1]),exp(p[2])))^2)/length(S)
  }

  SS_bevHolt <- function(p,S,R){
    sum((log(R)-bevHolt(S,exp(p[1]),exp(p[2])))^2)/length(S)
  }

  fitHockeyStick <- function(S,R){
    keepVals <- is.finite(S) & is.finite(R)
    S <- S[keepVals]
    R <- R[keepVals]
    p <- c(max(log((R+0.001)/(S+0.001))),mean(log(R)))
    tmp <- optim(p,SS_HockeyStick, method = "SANN",control = list(temp=5000,tmax=5000),S=S,R=R)
    tmp <- nlm(SS_HockeyStick,tmp$par ,S=S,R=R)
    tmp$par <- c(tmp$estimate,tmp$minimum)
    tmp
  }

  fitBevHolt <- function(S,R){
    keepVals <- is.finite(S) & is.finite(R)
    S <- S[keepVals]
    R <- R[keepVals]
    p <- c(max(log((R+0.001)/(S+0.001))),mean(log(R)))
    tmp <- optim(p,SS_bevHolt, method = "SANN",control = list(temp=5000,tmax=5000),S=S,R=R)
    tmp <- nlm(SS_bevHolt,tmp$par ,S=S,R=R)
    tmp$par <- c(tmp$estimate,tmp$minimum)
    tmp
  }

  addBevHolt <- function(S,R){
    tmp <- fitBevHolt(S,R)
    xx <- seq(from=0,max(S,na.rm=T),length.out=100)
    lines(xx,exp(bevHolt(xx,exp(tmp$par[1]),exp(tmp$par[2]))),lwd=2,col=rgb(0.1,0.1,0.9,0.5))
    invisible(tmp)
  }

  addHockeyStick <- function(S,R){
    tmp <- fitHockeyStick(S,R)
    xx <- seq(from=0,max(S,na.rm=T),length.out=100)
    lines(xx,exp(hockeyStick(xx,exp(tmp$par[1]),exp(tmp$par[2]))),lwd=2,col=rgb(0.1,0.1,0.9,0.5))
    invisible(tmp)
  }


```

```{r}
sites <- unique(tDat$Site)
par(mfrow=c(2,3))

for(site in sites){
  tt <- tDat[tDat$Site==site,]
  S <- tt$femaleSpawners/tt$area
  R <- tt$Smolts/tt$area
  plot(S,R,pch=16,xlim=c(0,max(S,na.rm=TRUE)),ylim=c(0,max(R,na.rm=TRUE)),bty="l",main=site)
  bhVals <- addBevHolt(S,R)
  hsVals <- addHockeyStick(S,R)
}
```

All together standardizing by basin area.

```{r}
S <- tDat$femaleSpawners/tDat$area
R <- tDat$Smolts/tDat$area
library(RColorBrewer)
pCols <- brewer.pal(n = 6, name = "Dark2")
names(pCols) <- unique(tDat$Site)
plot(S,R,pch=16,xlim=c(0,max(S,na.rm=TRUE)*1.1),
     ylim=c(0,max(R,na.rm=TRUE)*1.1),col=pCols[tDat$Site],
     bty="l",xaxs="i",yaxs="i")
legend("bottomright",legend=names(pCols),col=pCols,pch=rep(16,6))
```

Same thing but standardizing by miles

```{r}
S <- tDat$femaleSpawners/tDat$miles
R <- tDat$Smolts/tDat$miles
library(RColorBrewer)
pCols <- brewer.pal(n = 6, name = "Dark2")
names(pCols) <- unique(tDat$Site)
plot(S,R,pch=16,xlim=c(0,max(S,na.rm=TRUE)*1.1),
     ylim=c(0,max(R,na.rm=TRUE)*1.1),col=pCols[tDat$Site],
     bty="l",xaxs="i",yaxs="i")
legend("bottomright",legend=names(pCols),col=pCols,pch=rep(16,6))
```


```{r}
datLCMall <- createJAGSdata(dataDir,"smolt",habType="miles")
datOCNall <- createJAGSdata(dataDir,"OCN",habType="miles")
datLCM <- datLCMall$jagsDat
datOCN <- datOCNall$jagsDat

firstYrLCM <- min(createJAGSdata(dataDir,"smolt")$fdat$BroodYear)
firstYrOCN <- min(createJAGSdata(dataDir,"OCN")$fdat$broodYear)
yrDiff <- (firstYrLCM-firstYrOCN)

# stack the LCM data ontop of the OCN data
dat <- list()

# data for all years and populations (process model N total)
dat$N <- datLCM$N + datOCN$N
dat$stock <- c(datLCM$stock, datOCN$stock+datLCM$Npops)
dat$LCM <- rep(c(1,0), c(datLCM$N,datOCN$N))
dat$pHOS <- c(datLCM$pHOS,datOCN$pHOS)
dat$HR <- c(datLCM$HR,datOCN$HR)
dat$year <- c(datLCM$year + yrDiff,datOCN$year)
dat$escStartInd <- c(datLCM$escStartInd, datOCN$escStartInd + datLCM$N)
dat$escStopInd <- c(datLCM$escStopInd, datOCN$escStopInd + datLCM$N)
dat$Nyears <- diff(range(dat$year))+1

# populations
dat$Npops <- datLCM$Npops + datOCN$Npops
dat$habVar <- c(datLCM$habVar,datOCN$habVar)

# observed escapement
dat$escapementObs <- c(datLCM$escapementObs,datOCN$escapementObs)
dat$escInd <- c(datLCM$escInd,datOCN$escInd+datLCM$Nesc)
dat$Nesc <- datLCM$Nesc + datOCN$Nesc

# observed smolt (only for LCM pops)
dat$smoltObs <- datLCM$smoltObs
dat$smoltInd <- datLCM$smoltInd
dat$Nsmolt <- datLCM$Nsmolt

popNames <- c(datLCMall$siteNames,datOCNall$siteNames)
year <- c(datLCMall$fdat$BroodYear,datOCNall$fdat$broodYear)
years <- min(year):max(year)
```

```{r}
library(ggplot2)

yrInd <- which((years %% 5)==0)
yrLabs <- years[yrInd]

tDat <- data.frame(
  Escapement = rep(NA,dat$N),
  Smolt = rep(NA,dat$N),
  Population = dat$stock,
  Year = dat$year,
  pHOS = dat$pHOS,
  HR = dat$HR,
  LCM = dat$LCM,
  oSurv = rep(NA,dat$N),
  A2A = rep(NA,dat$N)
)
tDat$Escapement[dat$escInd] <- dat$escapementObs
tDat$Smolt[dat$smoltInd] <- dat$smoltObs

for(p in 1:dat$Npops){
  e1 <- dat$escStartInd[p]
  e2 <- dat$escStopInd[p]
  eRng <- (e1+3):e2
  if(dat$LCM[p]==1){
    tDat$oSurv[eRng] <- tDat$Escapement[eRng]/(1-tDat$HR[eRng])/tDat$Smolt[eRng-3]
  }
  tDat$A2A[eRng] <- (tDat$Escapement[eRng]/(1-tDat$HR[eRng]))/
                    (tDat$Escapement[eRng-3]/(1-tDat$pHOS[eRng-3]))
}

tmpNams <- popNames
names(tmpNams)=1:dat$Npops
pop.names <- as_labeller(tmpNams)

# ggplot(tDat,aes(x=Year,y=Escapement)) +
#   geom_line() +
#   facet_grid(Population~.,scales="free_y")
# 
# ggplot(tDat[tDat$LCM==1,],aes(x=Year,y=Smolt)) +
#   geom_line() +
#   facet_grid(Population~.,scales="free_y")
# 
# ggplot(tDat[tDat$LCM==1,],aes(x=Year,y=Smolt/Escapement)) +
#   geom_line() +
#   facet_grid(Population~.,scales="free_y")
# 
# ggplot(tDat[tDat$LCM==1,],aes(x=Year,y=oSurv)) +
#   geom_line() +
#   facet_grid(Population~.,scales="free_y",labeller=pop.names)

tmp <- tDat %>% group_by(Population) %>%
  mutate(escStnd=as.numeric(scale(log(Escapement))),
         smoltStnd=as.numeric(scale(log(Smolt))),
         A2AStnd=as.numeric(scale(log(A2A))),
         oSurvStnd=as.numeric(scale(log(oSurv))))%>%
  ungroup()
```

Here's a plot of log standardized escapement (orange is negative and blue positive)

```{r fig.width=8,fig.height=12}
pCol = c("orange","slateblue")
ggplot(tmp,aes(x=Year,y=Population)) +
  geom_point(aes(size=abs(tmp$escStnd)),color=pCol[sign(tmp$escStnd)/2+1.5]) +
  scale_size_continuous(range = c(0,4)) +
  scale_x_continuous(breaks=yrInd, labels=yrLabs) +
  scale_y_continuous(breaks = 1:dat$Npops,labels=popNames) +
  theme(axis.text.x = element_text(color="black", size=7, angle=45)) +
  theme_bw()
```

Same thing for (adult[y+3]/(1-HR[y+3])) / (adult[y]/(1-pHOS[y]))

```{r fig.width=8,fig.height=12}
ggplot(tmp,aes(x=Year,y=Population)) +
  geom_point(aes(size=abs(tmp$A2AStnd)),color=pCol[sign(tmp$A2AStnd)/2+1.5]) +
  scale_size_continuous(range = c(0,4)) +
  scale_x_continuous(breaks=yrInd, labels=yrLabs) +
  scale_y_continuous(breaks = 1:dat$Npops,labels=popNames) +
  theme(axis.text.x = element_text(color="black", size=7, angle=45)) +
  theme_bw()
```

Same thing for smolt to adult survival.
(adult[y+3]/(1-HR[y+3])) / smolt[y]

```{r fig.width=8,fig.height=6}
lcmDat <- tmp[tmp$LCM==1,]
ggplot(lcmDat,aes(x=Year,y=Population)) +
  geom_point(aes(size=abs(lcmDat$oSurvStnd)),color=pCol[sign(lcmDat$oSurvStnd)/2+1.5]) +
  scale_size_continuous(range = c(0,4)) +
  scale_x_continuous(breaks=yrInd, labels=yrLabs) +
  scale_y_continuous(breaks = 1:dat$Npops,labels=popNames) +
  theme(axis.text.x = element_text(color="black", size=7, angle=45)) +
  theme_bw()
```

Now let's use some back of the envelope calculations to estimate some parameters.

```{r}
  hockeyStick <- function(S,a,b){
    pmin(log(a)+log(S),log(b))
  }

  bevHolt <- function(S,a,b){
    log(S/(1/a+S/b))
  }

  SS_HockeyStick <- function(p,S,R){
    sum((log(R)-hockeyStick(S,exp(p[1]),exp(p[2])))^2)
  }

  SS_bevHolt <- function(p,S,R){
    sum((log(R)-bevHolt(S,exp(p[1]),exp(p[2])))^2)
  }

  fitHockeyStick <- function(S,R){
    keepVals <- is.finite(S) & is.finite(R)
    S <- S[keepVals]
    R <- R[keepVals]
    p <- c(max(log((R+0.001)/(S+0.001))),mean(log(R)))
    tmp <- optim(p,SS_HockeyStick, method = "SANN",control = list(temp=5000,tmax=5000),S=S,R=R)
    tmp <- nlm(SS_HockeyStick,tmp$par ,S=S,R=R)
    tmp$par <- tmp$estimate
    tmp
  }

  fitBevHolt <- function(S,R){
    keepVals <- is.finite(S) & is.finite(R)
    S <- S[keepVals]
    R <- R[keepVals]
    p <- c(max(log((R+0.001)/(S+0.001))),mean(log(R)))
    tmp <- optim(p,SS_bevHolt, method = "SANN",control = list(temp=5000,tmax=5000),S=S,R=R)
    tmp <- nlm(SS_bevHolt,tmp$par ,S=S,R=R)
    tmp$par <- tmp$estimate
    tmp
  }

  addBevHolt <- function(S,R){
    tmp <- fitBevHolt(S,R)
    xx <- seq(from=0,max(S,na.rm=T),length.out=100)
    lines(xx,exp(bevHolt(xx,exp(tmp$par[1]),exp(tmp$par[2]))),lwd=2,col=rgb(0.1,0.1,0.9,0.5))
  }

  addHockeyStick <- function(S,R){
    tmp <- fitHockeyStick(S,R)
    xx <- seq(from=0,max(S,na.rm=T),length.out=100)
    lines(xx,exp(hockeyStick(xx,exp(tmp$par[1]),exp(tmp$par[2]))),lwd=2,col=rgb(0.1,0.1,0.9,0.5))# rgb(0.9,0.6,0.1,0.5))
  }


```


