---
title: "OCN coho"
author: "Martin Liermann"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: sandstone
    highlight: tango 
    toc: true
    toc_float: true
    code_folding: hide
params:
  runMod: FALSE
---

```{r}
# center drive working directory
workDir <- "//nwcfile/home/liermannma/CurrentWorrk/consulting/sept2018_sept2019/oregonCoho"

# local computer working directory
workDir <-"C:\\Users\\Martin.Liermann\\Documents\\projects\\oregonCoho"

# directory with the data and run results
dataDir <- paste(workDir,"data",sep="/")
runsDir <- paste(workDir,"runs",sep="/")

# load the library 
library(coastalCohoSS)

# define some colors
greenT <- rgb(0.2,0.8,0.2,0.2)

```

```{r}
# data from Bradford 1995
#  these are egg to smolt instananeous mortality rates
#  so survival is exp(-M)
sDat <- data.frame(
  population=c("Black","Deer","Flynn","Carnation","Hunt",
               "Karymaisky","Needle","Minter","Nile","Qualicum"),
  mean=c(4.17,3.37,4.02,3.88,4.64,5.93,4.45,3.79,4.33,4.40),
  sd=c(0.93,0.41,0.90,0.47,1.04,0.77,0.89,0.83,0.42,0.79)
  )
sDat$surv <- exp(-sDat$mean)
qM <- quantile(sDat$mean,prob=c(0,0.25,0.5,0.75,1.0))
qS <- quantile(sDat$surv,prob=c(0,0.25,0.5,0.75,1.0))
```

Run the model for the 21 OCN populations using the Beverton Holt, modified Beverton Holt and Hockey stick. 

```{r eval=params$runMod}
MCMCsims <- 100000

library(coastalCohoSS)
dat <- createJAGSdata(dataDir,"OCN")
initValFunc <- createInitValsFunc(dat)
priors <- createDefaultPriors()
priors$prodMuPior <- c(log(4),1/(2^2))

bmod <- createJAGScode(fixedObsError=c(0.15,0.15), smoltData=FALSE,SRtype="HS")
m1 <- runJAGSmodel(bmod, dat, priors, calcInits=initValFunc, MCMCsims=MCMCsims)
bmod <- createJAGScode(fixedObsError=c(0.15,0.15), smoltData=FALSE,SRtype="BH")
m2 <- runJAGSmodel(bmod, dat, priors, calcInits=initValFunc, MCMCsims=MCMCsims)
bmod <- createJAGScode(fixedObsError=c(0.15,0.15), smoltData=FALSE,SRtype="BH3")
m3 <- runJAGSmodel(bmod, dat, priors, calcInits=initValFunc, MCMCsims=MCMCsims)

saveRDS(m1,paste(runsDir,"mHSocn.Rdat",sep="/"))
saveRDS(m2,paste(runsDir,"mBHocn.Rdat",sep="/"))
saveRDS(m3,paste(runsDir,"mBH3ocn.Rdat",sep="/"))
```
+
```{r eval=!params$runMod}
m1 <- readRDS(paste(runsDir,"mHSocn.Rdat",sep="/"))
m2 <- readRDS(paste(runsDir,"mBHocn.Rdat",sep="/"))
m3 <- readRDS(paste(runsDir,"mBH3ocn.Rdat",sep="/"))
```

```{r}
# calculate some intermediate values
x1 <- getPostDraws(m1)
x2 <- getPostDraws(m2)
x3 <- getPostDraws(m3)
```

Look at productivity estimates.

```{r}
# a function to create a long form of the posterior distribution for
# parameters that are population specific (e.g. productivity and capacity)
createLong <- function(x,pVar){
  tmp <- as.data.frame(x[[pVar]])
  colnames(tmp) <- m1$dat$siteNames
  ss <- tmp %>% pivot_longer(everything(),names_to="River")
}

# combine the results for the different spawner recruit functions
SRfuncNames <- c("Hockey stick","Beverton Holt","Beverton Holt (mod)")
xx <- rbind(createLong(x1,"prod"),createLong(x2,"prod"),createLong(x3,"prod"))
n <- length(xx$River)/3
xx$SRfunction <- rep(SRfuncNames,each=n)

# create results for a "new" population with no data
xx2 <- c(x1$logProdMu,x2$logProdMu,x3$logProdMu)
n <- length(xx2)/3
xx2 <- data.frame(SRfunction=rep(SRfuncNames,each=n),logProdMu=xx2)
xx2$logProdSD <- c(x1$logProdSD,x2$logProdSD,x3$logProdSD)
xx2$newProd <- rnorm(n*3,xx2$logProdMu,xx2$logProdSD)

# create a data frame with median productivity values
xx3 <- data.frame(medianProd=c(apply(x1$prod,2,median),apply(x2$prod,2,median),apply(x3$prod,2,median)))
n <- length(xx3$medianProd)/3
xx3$River <- rep(m1$dat$siteNames,3)
xx3$SRfunction <- rep(SRfuncNames,each=n)

```

```{r}
# bar plot summarizing productivity for the different populations and SR types
brks <- 2^(0:50)
ggplot(xx,aes(x=River,y=value,fill=SRfunction)) + 
  geom_boxplot(color="black",outlier.shape = NA) +
  coord_flip(ylim=c(1,2024)) +
  theme_bw() +
  scale_y_continuous(trans='log2',breaks=brks,labels=paste(brks))
```

```{r}
# plot describing the distribution for a new population
brks <- 2^(0:50) 
ggplot(xx2,aes(x=exp(newProd),fill=SRfunction,color=SRfunction)) +
  geom_density(alpha = 0.1) +
  theme_bw() +
  scale_x_continuous(trans='log2',breaks=brks,labels=paste(brks)) +
  coord_cartesian(xlim=c(1,1000))+
  facet_grid(SRfunction~.) + 
  geom_vline(data=xx3,aes(xintercept=medianProd))
```

```{r message=FALSE}
# Table with results
qProd <- xx %>% 
  group_by(River,SRfunction) %>%
  summarize(med=median(value),q25=quantile(value,prob=0.25),q75=quantile(value,prob=0.75))

qProdNew <- xx2 %>% 
  group_by(SRfunction) %>%
  mutate(value=exp(newProd)) %>%
  summarize(med=median(value),
            q25=quantile(value,prob=0.25),
            q75=quantile(value,prob=0.75),
            muLogProd=mean(newProd),sdLogProd=sd(newProd))

knitr::kable(qProdNew,digits=3)
```

```{r}
# temporal trends in common year effect
x <- x3
dat <- m1$dat
yeQ <- apply(x$yearEffect,2,quantile,prob=c(0.1,0.25,0.5,0.75,0.9))
yRange <- c(log(1/4),log(4))
yrs <- seq(min(dat$fdat$broodYear),max(dat$fdat$broodYear),by=1)
lineCol <- rgb(0.99,0.64,0.0,0.25)

plot(yrs,yeQ[3,],ylim=yRange,xlab="Brood Year",ylab="Common year effect",type="o",bty="l",yaxt="n")
lines(yrs,yeQ[2,],lty=3)
lines(yrs,yeQ[4,],lty=3)
pSamp <- sample(1:length(x$prod[,1]),25)
for(smp in pSamp){
  lines(yrs,x$yearEffect[smp,],col=lineCol)
}

yLabs <- 2^((-4:4))
axis(side=2,at=log(yLabs),labels=yLabs,las=2)

for(yy in yLabs) lines(range(yrs),rep(log(yy),2),lty=3,col="gray")

legend("bottomright",lty=c(1,3),legend=c("Median","50% credible interval"))

```

```{r}
# temporal trends in productivity
x <- x3
dat <- m1$dat
yeQ <- apply(x$yearEffect+as.vector(x$logProdMu),2,quantile,prob=c(0.1,0.25,0.5,0.75,0.9))
yRange <- c(log(1),log(150))
yrs <- seq(min(dat$fdat$broodYear),max(dat$fdat$broodYear),by=1)
lineCol <- rgb(0.99,0.64,0.0,0.25)

plot(yrs,yeQ[3,],ylim=yRange,xlab="Brood Year",ylab="Common year effect",type="o",bty="l",yaxt="n")
lines(yrs,yeQ[2,],lty=3)
lines(yrs,yeQ[4,],lty=3)
pSamp <- sample(1:length(x$prod[,1]),25)
for(smp in pSamp){
  lines(yrs,(x$yearEffect+as.vector(x$logProdMu))[smp,],col=lineCol)
}

yLabs <- 2^((-4:20))
axis(side=2,at=log(yLabs),labels=yLabs,las=2)

for(yy in yLabs) lines(range(yrs),rep(log(yy),2),lty=3,col="gray")

legend("bottomright",lty=c(1,3),legend=c("Median","50% credible interval"))

```

```{r}
# some tabular values
yrs <- min(m1$dat$fdat$broodYear):max(m1$dat$fdat$broodYear)
tmpDat <- data.frame(broodYear=yrs,exp(t(yeQ)))
names(tmpDat) <- c("BroodYear","q10","q25","q50","q75","q90")
knitr::kable(tmpDat,digits=0)
```

```{r}
# temporal trends in productivity

```




```{r}
# extra stuff
brks <- 2^(0:50) 
ggplot(xx2,aes(x=exp(newProd),fill=SRfunction,color=SRfunction)) +
  geom_density(alpha = 0.1) +
  geom_density(alpha = 0.1) +
  theme_bw() +
  scale_x_continuous(trans='log2',breaks=brks,labels=paste(brks)) +
  coord_cartesian(xlim=c(1,1000))+
  facet_grid(SRfunction~.) + 
  geom_vline(data=xx3,aes(xintercept=medianProd))

brks <- 2^(0:50) 
ggplot(xx2,aes(x=exp(newProd),fill=SRfunction,color=SRfunction)) +
  geom_density(alpha = 0.1) +
  theme_bw() +
  scale_x_continuous(trans='log2',breaks=brks,labels=paste(brks)) +
  coord_cartesian(xlim=c(1,1000))+
  facet_grid(SRfunction~.) + 
  geom_boxplot(data=xx3,aes(xintercept=medianProd))


ggplot(xx2,aes(x=SRfunction,y=exp(newProd),fill=SRfunction)) + 
  geom_boxplot(color="black",outlier.shape = NA) +
  coord_flip() +
  theme_bw() +
  scale_y_continuous(trans='log2',breaks=brks,labels=paste(brks))

ggplot(xx2,aes(x=SRfunction,y=exp(newProd),fill=SRfunction)) +
  geom_boxplot(color="black",outlier.shape = NA) +
  coord_flip() +
  theme_bw() +
  scale_y_continuous(trans='log2',breaks=brks,labels=paste(brks)) +
  coord_cartesian(ylim = c(0,20))
 
brks <- 2^(0:50) 
ggplot(xx2,aes(x=SRfunction,y=exp(newProd),fill=SRfunction)) +
  geom_boxplot(color="black",outlier.shape = NA) +
  theme_bw() +
  scale_y_continuous(trans='log2',breaks=brks,labels=paste(brks)) +
  coord_flip(ylim=c(1,1000)) 

brks <- 2^(0:50) 
ggplot(xx2,aes(x=exp(newProd),fill=SRfunction,color=SRfunction)) +
  geom_density(alpha = 0.1) +
  theme_bw() +
  scale_x_continuous(trans='log2',breaks=brks,labels=paste(brks)) +
  coord_cartesian(xlim=c(1,1000))+
  facet_grid(SRfunction~.)
```




```{r fig.width=12, fig.height=12, warning=FALSE}
# create a plot of spawner to adult recruits (for 6 pops)
#  and overlay the two model fits (use two bands)

bdat <- m1$dat$jagsDat
edat <- expandAllData(bdat)
popNames <- m1$dat$siteNames
sims <- length(x1$prod[,2])
Npops <- bdat$Npops

# calculate the the productivity and capacity values for the spawner to spawner SR functions
OS <- array(rep(tapply(m3$dat$jagsDat$oceanSurv,m3$dat$jagsDat$stock,median),each=sims),
            dim=c(sims,Npops))
prodWOS <- x2$prod*OS
capWOS <- x2$cap*OS

# calculate different ocean regimes

#####  spawners to trap model (prepare values) #####
spawnersNO <- edat$escapementObs # natural origin spawners
spawners <- spawnersNO/(1-edat$pHOS)
Rtemp <- spawnersNO/(1-edat$HR)
recruits <- numeric(length(Rtemp))
for(i in 1:bdat$Npops){
  recruits[bdat$stock==i] <- c(Rtemp[bdat$stock==i][-(1:3)],rep(NA,3))
}

##### estimate the SR params #####
srF <- function(p,ss) ss/(1/p[1]^3 + ss^3/p[2]^3)^(1/3)
fitBH <- function(ss,rr){
  ssq <- function(pp,ss,rr){
    p <- exp(pp)
    sum((log(rr)-log(srF(p,ss)))^2)
  }
  bFit <- nlm(ssq,p=log(c(median(rr/ss),quantile(rr,prob=0.8))),ss,rr)
  bFit
}

srDat <- data.frame(S=spawners, R=recruits, stock=edat$stock)
srDat <- srDat[!is.na(srDat$S) & !is.na(srDat$R),] # remove NAs
srParams <- array(NA,dim=c(bdat$Npops,2))

par(mfrow=c(5,5))

addBand <- function(prod,cap,ss,col=rgb(0,0,0,0.2)){
  tmp <- col2rgb(col)
  lnCol <- rgb(tmp[1]/255,tmp[2]/255,tmp[3]/255,0.9)
  n <- length(ss)
  R <- array(NA,dim=c(3,n))
  for(i in 1:n){
    R[,i] <- quantile(ss[i]/(1/prod^3+(ss[i]/cap)^3)^(1/3),prob=c(0.1,0.5,0.9))
  }
  polygon(x=c(ss,rev(ss)),y=c(R[1,],rev(R[3,])),col=col,border=rgb(0,0,0,0.5))
  lines(ss,R[2,],col=lnCol)
}

for(st in 1:bdat$Npops){
  sInd <- srDat$stock==st
  ss <- seq(0,max(srDat$S[sInd]),length.out=1001)
  xLim <- c(0,max(srDat$S[sInd]))
  yLim <- c(0,max(srDat$R[sInd]))
  r1 <- fitBH(srDat$S[sInd],srDat$R[sInd])
  srParams[st,] <- exp(r1$estimate)
  plot(srDat$S[sInd],srDat$R[sInd],pch=16,col="black",xlab="Spawners",ylab="Recruits",bty="l",main=popNames[st],xlim=xLim,ylim=yLim)
  lines(ss,srF(exp(r1$estimate),ss=ss),lwd=2,col="black")
  addBand(prodWOS[,st],capWOS[,st],ss,rgb(0.2,0.2,0.8,0.2))
  grid()
}
